\documentclass[nopalatino,nolot,nolof,color,11pt]{fithesis3}
\thesissetup{
  faculty=fi,
  type=mgr,
  date={2015/12/31},
  author={Bc. Pavel Dedík},
  advisor={doc. Mgr. Radek Pelánek, Ph.D.},
  title={Modeling Human Memory for Adaptive Educational Systems},
  keywords={Memory, Learning, Forgetting, Knowledge, Adaptive Educational System, Machine Learning, Student Modeling, Performance Factor Analysis},
  abstract={Adaptive educational systems enable students to practice different domains of educational content adaptively and make learning personalized. In this thesis, we study and design extensions of student models suitable for applications in adaptive systems while focusing on key aspects of human memory and learning. We evaluate the proposed models using data from the adaptive system for practice of geography facts available at \href{http://outlinemaps.org}{\texttt{outlinemaps.org}}, and compare currently used models with ours. Finally, we analyze parameter values of the models and discuss the results.},
  thanks={Firstly, I would like to express my gratitude to my advisor, Radek~Pelánek, for his patience, motivation and helpful guidance. My thanks also goes to members of the Adaptive Learning group for their comments and remarks. I~am also grateful to my brother for his stylistic revision, and my friends who provided some feedback.}
}

\usepackage{color}
\usepackage{lmodern}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{wrapfig}
\usepackage{footnote}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{appendix}
\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{mathtools}
\usepackage{subcaption}
\usepackage{threeparttable}
\usepackage{etoolbox,siunitx}
\usepackage{booktabs,multirow}
\usepackage[noend]{algpseudocode}
\usepackage[labelfont=bf]{caption}

\setlist[itemize]{leftmargin=10mm}

\begin{document}
  \input{introduction}
  \input{background}
  \input{models}
  \input{analysis}
  \input{discussion}
  \input{conclusion}

  \bibliography{references}
  \bibliographystyle{IEEEtran}

  \begin{appendices}
  \input{appendix}
  \end{appendices}
\end{document}
